---
bibliography: references.bib
---

```{=tex}
\vspace*{-3\baselineskip}
\section{\textgreek{η}/\textgreek{ᾱ}}
```

```{r Find_h_data}
# h / a: ------------------------------------------------------------
## DHMOS ------------------------------------------------------------
# Read in assigned data
Data.h.dhmos <- read_csv(
  "~/Documents/UGA/Thesis/Wukia/data/Data.h.dhmos.CONF.csv",
  col_types = cols(
    ig_book = col_factor(levels = c()),
    ig_no = col_factor(levels = c()),
    location = col_factor(levels = c()),
    token = col_factor(levels = c()),
    data_type = col_factor(levels = c()),
    simplex = col_logical()
  )
)

# Remove erroneous captures
Data.h.dhmos <- Data.h.dhmos[!is.na(Data.h.dhmos$data_type), ]
Data.h.dhmos <- Data.h.dhmos[!is.na(Data.h.dhmos$simplex), ]


# I... think I need to drop Δημήτριος
Data.h.dhmos <- Data.h.dhmos %>%
  filter(str_detect(token, "(δημητρ|δαματρ)") == F)

# Drop unused levels
Data.h.dhmos <- droplevels(Data.h.dhmos)

## GH ---------------------------------------------------------------
# Read in assigned data
Data.h.gh <- read_csv(
  "~/Documents/UGA/Thesis/Wukia/data/Data.h.gh.CONF.csv",
  col_types = cols(
    ig_book = col_factor(levels = c()),
    ig_no = col_factor(levels = c()),
    location = col_factor(levels = c()),
    token = col_factor(levels = c()),
    data_type = col_factor(levels = c()),
    simplex = col_logical()
  )
)

# Remove erroneous captures
Data.h.gh <- Data.h.gh[!is.na(Data.h.gh$data_type), ]

# Drop unused levels
Data.h.gh <- droplevels(Data.h.gh)

## BOULH ------------------------------------------------------------
# Read in assigned data
Data.h.boulh <- read_csv(
  "~/Documents/UGA/Thesis/Wukia/data/Data.h.boulh.CONF.csv",
  col_types = cols(
    ig_book = col_factor(levels = c()),
    ig_no = col_factor(levels = c()),
    location = col_factor(levels = c()),
    token = col_factor(levels = c()),
    data_type = col_factor(levels = c()),
    simplex = col_logical()
  )
)

# Remove erroneous captures
Data.h.boulh <- Data.h.boulh[!is.na(Data.h.boulh$data_type), ]

# Drop unused levels
Data.h.boulh <- droplevels(Data.h.boulh)

## STHLH ------------------------------------------------------------
# Read in assigned data
Data.h.sthlh <- read_csv(
  "~/Documents/UGA/Thesis/Wukia/data/Data.h.sthlh.CONF.csv",
  col_types = cols(
    ig_book = col_factor(levels = c()),
    ig_no = col_factor(levels = c()),
    location = col_factor(levels = c()),
    token = col_factor(levels = c()),
    data_type = col_factor(levels = c()),
    simplex = col_logical()
  )
)

# Remove erroneous captures
Data.h.sthlh <- Data.h.sthlh[!is.na(Data.h.sthlh$data_type), ]

# Drop unused levels
Data.h.sthlh <- droplevels(Data.h.sthlh)

## COMBINED ---------------------------------------------------------
# Combine all assigned data tables
Data.h <- Data.h.sthlh %>%
  full_join(Data.h.gh) %>%
  full_join(Data.h.dhmos) %>%
  full_join(Data.h.boulh)
```

The \textgreek{η}/\textgreek{ᾱ} isogloss is evidenced by the vowel of the root syllables in \textgreek{δῆμος} 'people, assembly' and \textgreek{στήλη} 'stele,' as well as the \textipa{\=a}-stem endings in \textgreek{βουλή} 'council' and \textgreek{γῆ} 'land.'

First, the noun \textgreek{δᾶμος} occurs 27 times in the exemplary corpus, often in resolution formulas indicating that a decree had passed all legal prerequisites for enactment, such as in DGE 459.2 \textgreek{δεδόχθη} \textgreek{τοῖ} \textgreek{δάμοι} 'be it resolved by the people.'[@mclean2011, 231] Next, the noun \textgreek{στάλα} appears 7 times, each in a formula determining the method of public presentation for a decree. For example, see the dative \textgreek{στάλαι} in DGE 155: \textgreek{ἀναγραψάτω δὲ τὸ δόγμα τόδε ὁ γραμματεὺς τοῦ δάμου ἐν στάλαι λιθίναι καὶ ἀνθέτω εἰς τὸ Ὀλυμπιεῖον} 'let the secretary of the people engrave this edict on a stone stele and set it up at the Olympieion.'

Additionally, the noun \textgreek{βουλά} occurs 8 times in the exemplary data. For example, the genitive singular \textgreek{βουλᾶς} appears in DGE 155 in an opening formula listing the standing city officials in order to date the decree: \textgreek{γραμματεὺς βουλᾶς καὶ δάμου Ἵππων Παγχάρεος} 'Hippon son of Panchareus (was the) Secretary of the council and the assembly.' Finally, the noun \textgreek{γᾶ} appears 15 times in the exemplary corpus, typically in proxeny decrees issuing rights and protections to public friends, as seen in DGE 460.1 \textgreek{γᾶς κὴ ϝυκίας ἔππασιν κὴ ϝισοτελίαν} 'the right to purchase land and a house, as well as equality in taxes.'

However, the exemplary data also includes 1 occurrence of the \textgreek{η}-type genitive \textgreek{γῆς} in a similar formula from the 4\super{th} cent. Oropian inscription DGE 812: \textgreek{καὶ γῆς καὶ οἰκίης ἔνκτησιν αὐτῶι καὶ ἐκγόνοις} 'the right to purchase land and a house for him and his descendants.' Additionally, the contemporary inscription DGE 811, also from Oropos, contains 3 instances of \textgreek{δῆμος} and its derivatives, such as \textgreek{δημότης} 'citizen' in the phrase \textgreek{ἂν δέ τις ἀδικεῖ ἐν τοῖ ἱεροῖ ἢ ξένος ἢ δημότης} 'if one does wrong in the temple, be they either a foreigner or a citizen.' The presence of these \textgreek{η}-type tokens brings into question the full geographical extent of the \textgreek{ᾱ}\>\textgreek{η} vowel shift.

### 4\super{th} cent. BCE

```{r Load_4th_cen_h_data}
### City Counts -----------------------------------------------------
Data.h.4 <-
  Data.h[Data.h$analysis_cen == 400, ] %>% # . . oi for 4th cen
  droplevels()

Count.h.4 <- Data.h.4 %>% nrow()
Count.h.4.Athens <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>% nrow()
Count.h.4.Eleusis <- Data.h.4 %>%
  filter(str_detect(location, "Eleusis") == T) %>% nrow()
Count.h.4.Oropos <- Data.h.4 %>%
  filter(str_detect(location, "Oropos") == T) %>% nrow()
Count.h.4.Megara <- Data.h.4 %>%
  filter(str_detect(location, "Megara") == T) %>% nrow()
Count.h.4.Tanagra <- Data.h.4 %>%
  filter(str_detect(location, "Tanagra") == T) %>% nrow()
```

The data from the 4\super{th} cent. BCE establishes the \textgreek{η}/\textgreek{ᾱ} isogloss as dividing Attica and Oropia from Megaris and Boiotia. The set contains `r Count.h.4` tokens from 5 cities: Athens (n=`r Count.h.4.Athens`), Eleusis (n=`r Count.h.4.Eleusis`), Oropos (n=`r Count.h.4.Oropos`), Megara (n=`r Count.h.4.Megara`), and Tanagra (n=`r Count.h.4.Tanagra`). The \textgreek{η}-type and \textgreek{ᾱ}-type tokens belong almost exclusively on their associated side of the isogloss line, with only 4 \textgreek{ᾱ}-type tokens appearing on the \textgreek{η} side.

```{r Visualize_4th_cen_h_tokens_by_city, fig.pos='p'}
#| label: fig_4th_cen_h_tokens
#| fig-cap: "Fourth Century BCE Token Counts by City"

# Visualize location counts
Data.h.4 %>%
  ggplot(aes(x = forcats::fct_infreq(location), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Location", y = "Token Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

#### Athens

`r Count.h.4.Athens` tokens represent the Athenian data. While the set does contain a small number of \textgreek{ᾱ}-type tokens, the data overall indicates that the Athenian population preferred forms with the innovative raised \textgreek{η} over those with \textgreek{ᾱ}.

First, the set contains `r Data.h.4 %>% filter(str_detect(location, "Athens") == T) %>% filter(str_detect(token, "δημ") == T) %>% nrow()` tokens built from \textgreek{δῆμος}. This list includes `r Data.h.4 %>% filter(str_detect(location, "Athens") == T) %>% filter(str_detect(token, "^δημωι$") == T) %>% nrow()` instances of the dative singular \textgreek{δήμωι} and `r Data.h.4 %>% filter(str_detect(location, "Athens") == T) %>% filter(str_detect(token, "^δημου$") == T) %>% nrow()` of the genitive \textgreek{δήμου}, as well as `r Data.h.4 %>% filter(str_detect(location, "Athens") == T) %>% filter(str_detect(token, "^δημον$") == T) %>% nrow()` of the accusative \textgreek{δήμον} and `r Data.h.4 %>% filter(str_detect(location, "Athens") == T) %>% filter(str_detect(token, "^δημος$") == T) %>% nrow()` of the nominative \textgreek{δῆμος}. The set further includes 80 compound tokens built with the member \textgreek{δημ}-.

```{r 4th_cen_sthl_counts_athens}
Data.h.4.sthl <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>%
  filter(str_detect(token, "στηλ") == T) %>%
  nrow()
Data.h.4.sthlhs <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>%
  filter(str_detect(token, "^στηλης$") == T) %>%
  nrow()
Data.h.4.sthlhi <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>%
  filter(str_detect(token, "^στηληι$") == T) %>%
  nrow()
Data.h.4.sthlas <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>%
  filter(str_detect(token, "^στηλας$") == T) %>%
  nrow()
```

Furthermore, the data contain `r Data.h.4.sthl` occurrences of \textgreek{στήλη}. Among these are `r Data.h.4.sthlhs` instances of the genitive singular \textgreek{στήλης}, as well as `r Data.h.4.sthlhi` each of the dative singular \textgreek{στήληι} and the accusative \textgreek{στήλην}, along with `r Data.h.4.sthlas` of the accusative plural \textgreek{στήλας} and 1 each of the dative \textgreek{στήλαις} and the nominative \textgreek{στήλη}.

```{r 4th_cen_boul_counts_athens}
Data.h.4.boul <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>%
  filter(str_detect(token, "βουλ") == T) %>%
  nrow()
Data.h.4.boulhs <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>%
  filter(str_detect(token, "^βουλης$") == T) %>%
  nrow()
Data.h.4.boulh.nom <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>% 
  filter(str_detect(token, "^βουλη$") == T) %>%
  nrow() 
Data.h.4.boulhi <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>%
  filter(str_detect(token, "^βουληι$") == T) %>%
  nrow()
Data.h.4.boulhn <- Data.h.4 %>%
  filter(str_detect(location, "Athens") == T) %>%
  filter(str_detect(token, "^βουλην$") == T) %>%
  nrow()
```

Besides these, the data include `r Data.h.4.boul` occurrences of \textgreek{βουλή}: `r Data.h.4.boulhs` being the genitive singular \textgreek{βουλῆς}, `r Data.h.4.boulh.nom` the nominative singular \textgreek{βουλή}, `r Data.h.4.boulhi` the dative \textgreek{βουλῆι}, and `r Data.h.4.boulhn` the accusative \textgreek{βουλήν}. And finally, the set includes 2 instances of \textgreek{γῆ} as the genitive singular \textgreek{γῆς}.

Therefore, with only \textgreek{η}-type tokens, the data characterize Athens as having accepted that \textgreek{ᾱ}\>\textgreek{η} shift and establish the city as the point of comparison for testing the position of other cities relative to the \textgreek{η}/\textgreek{ᾱ} isogloss line.

```{r 4th_cen_h_athen, fig.pos='p'}
#| label: fig_4th_h_athens
#| fig-cap: " Selected Fourth Century BCE Athenian Token Counts"

# Visualize tokens in Athens
Data.h.4[Data.h.4$location == "Athens",] %>%
  filter(simplex == 1) %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(subtitle = "Simplex Tokens",
       x = "Token",
       y = "Count",
       caption = "Chart displays 276 of 351 Athenian tokens.") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

#### Eleusis

```{r Load_4th_cen_eleusinean_h_data}
# Filter for Eleusinean data
Data.h.4.Eleusis <- Data.h.4 %>%
  filter(str_detect(location, "(Athens|Eleusis)") == T) %>%
  droplevels()
```

 The `r Count.h.4.Eleusis` tokens from Eleusis indicate that the city's population preferred forms with the raised \textgreek{η} over the older \textgreek{ᾱ}. The innovative \textgreek{η} occurs 80 times in the root \textgreek{δημ}-, including 13 instances of the nominative \textgreek{δῆμος}, 11 of the dative plural \textgreek{δημόσιοις} 'public,' and 9 of the genitive singular \textgreek{δήμου}. The data also includes various derived forms, such as the proper noun \textgreek{Χαρίδημος} 'Grace of the Deme' and \textgreek{Σωσίδημος} 'Savior of the Deme,' as well as the participle \textgreek{δημαρχῶν} 'being the Archon of the Deme.' Additionally, the root \textgreek{στηλ}- occurs twice: once as the genitive singular \textgreek{στήλης} and once as the dative singular \textgreek{στήληι}.

Furthermore, endings with -\textgreek{η}(-) occur 16 times appended with \textgreek{βουλή}. This includes 8 instances of the genitive singular \textgreek{βουλῆς}, 6 of the nominative \textgreek{βουλή}, 1 of the accusative \textgreek{βουλήν}, and 1 of the dative \textgreek{βουλῆι}. And finally, the data contain 3 instances of -\textgreek{η}(-) endings on the noun \textgreek{γῆ}. This includes 2 occurrences of the genitive singular \textgreek{γῆς} and 1 of the accusative singular \textgreek{γῆν}.

With no \textgreek{ᾱ}-type tokens in the set, the data categorize Eleusis as belonging to the Athenian side of the \textgreek{η}/\textgreek{ᾱ} isogloss line, indicating that the city's population participated in the Attic-Ionic vowel shift.

```{r 4th_cen_h_eleusis, fig.pos='t'}
#| label: fig_4th_h_eleusis
#| fig-cap: "Selected Fourth Century BCE Eleusinean Token Counts"

# Visualize tokens in Eleusis
Data.h.4[Data.h.4$location == "Eleusis",] %>%
  filter((simplex == 1) | (data_type == "α")) %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(subtitle = "Simplex and α-Type Tokens",
       x = "Token",
       y = "Count",
       caption = "Chart displays 57 of 105 Eleusinean tokens.") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

#### Oropos

The `r Count.h.4.Oropos` Oropian tokens unanimously indicate that the city participated in the same vowel shift. The data include 49 instances of \textgreek{η} within nouns. 41 appear in forms of \textgreek{δῆμος}, primarily in the 10 occurrences of the dative singular \textgreek{δήμωι}, 8 of the genitive \textgreek{δήμου}, and 5 of the accusative \textgreek{δήμον}. The last 8 appear in the word \textgreek{στήλη}: 4 times in the genitive \textgreek{στήλης}, 3 times in the dative \textgreek{στήληι}, and once in the accusative plural \textgreek{στήλας}.

In addition, the data include 15 \textgreek{η}-type tokens with endings in -\textgreek{η}(-): 10 forms of \textgreek{βουλή} and 5 of \textgreek{γῆ}. The \textgreek{βουλή} tokens are comprised of 7 instances of the dative singular \textgreek{βουλῆι} and 3 of the genitive \textgreek{βουλῆς}, whereas \textgreek{γῆ} tokens are comprised of 3 instances of the genitive \textgreek{γῆς} and 2 of the accusative \textgreek{γῆν}.

Overall, the `r Count.h.4.Oropos` tokens from Oropos all belong to the \textgreek{η}-type, indicating that the city belonged to the Athenian side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 4\super{th} cent. BCE.

```{r 4th_cen_h_oropos}
#| label: fig_4th_h_oropos
#| fig-cap: "Fourth Century BCE Oropian Token Counts"

# Visualize tokens in Oropos
Data.h.4[Data.h.4$location == "Oropos",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

#### Megara

```{r Load_4th_cen_megarian_h_data}
# Filter for Megarian data
Data.h.4.Megara <- Data.h.4 %>%
  filter(str_detect(location, "(Athens|Megara)") == T) %>%
  droplevels()
```

The `r Count.h.4.Megara` tokens from Megara show that the local population did not participate in the \textgreek{ᾱ}\>\textgreek{η} vowel shift. Every token shows the inherited long \textgreek{ᾱ} in both root syllables and \textipa{\=a}-stem endings.

First, the data contain 46 tokens with the root vowel \textgreek{ᾱ}. Among these are 39 instances of \textgreek{δᾶμος}, including, for example, the dative singular \textgreek{δάμωι} (n=14) and the genitive \textgreek{δάμου} (n=13). The stem also appears 7 times in proper nouns: 4 times in the nominative \textgreek{Δαμοτέλης} 'Fulfillment of the Deme' and 3 times in the patronymic genitive \textgreek{Δαμοτέλεος} 'Son of Damoteles.' In addition, the set includes 7 occurrences of the word \textgreek{στ\'ᾱλᾱ}: 4 of the accusative \textgreek{στάλαν} and 3 of the dative \textgreek{στάλαι}.

Furthermore, the Megarian data include 15 tokens with endings in -\textgreek{ᾱ}(-). The list is comprised of 7 instances of the dative singular \textgreek{βουλᾶι}, 6 of the genitive \textgreek{βουλᾶς}, and 2 of the genitive \textgreek{γᾶς}.

Finally, the set contains no \textgreek{η}-type tokens. Therefore, the data suggest uniform retention of the long \textgreek{ᾱ} in Megara. To corroborate this, the author performed Fisher's exact test on the Athenian and Megarian data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Megarian data of the 4\super{th} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Megarian data of the 4\super{th} cent. BCE.

```{r 4th_cen_h_megara, fig.pos='p'}
#| label: fig_4th_h_megara
#| fig-cap: "Fourth Century BCE Megarian Token Counts"

# Visualize tokens in Megara
Data.h.4[Data.h.4$location == "Megara",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r 4th_cen_h_tanagra, fig.pos='p'}
#| label: fig_4th_h_tanagra
#| fig-cap: "Fourth Century BCE Tanagran Token Counts"

# Visualize tokens in Tanagra
Data.h.4[Data.h.4$location == "Tanagra",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r 4th_cen_h_megara_fisher}
#| include: false

XT.h.4.Megara <-
  stats::xtabs( ~ data_type + location, data = Data.h.4.Megara)
# Fisher's Exact Test for Megarian Data
fisher.test(XT.h.4.Megara)
```

The test resulted in a *p*-value of \< 2.2 x 10^-16^. With such a small *p*-value, the author rejects the null hypothesis and asserts that, through the tokens collected, the data show a clear relationship between the origin of an inscription and the data type of the tokens therein. Therefore, Megara belonged on the \textgreek{ᾱ} side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 4\super{th} cent. BCE.

#### Tanagra

```{r Load_4th_cen_tanagran_h_data}
# Filter for Tanagran data
Data.h.4.Tanagra <- Data.h.4 %>%
  filter(str_detect(location, "(Athens|Tanagra)") == T) %>%
  droplevels()
```

The `r Count.h.4.Tanagra` Tanagran inscriptions show that the city did not participate in the \textgreek{ᾱ}\>\textgreek{η} vowel shift. First, The data include 3 instances of  \textgreek{δᾶμος}: once each as the accusative singular \textgreek{δᾶμον}, the dative \textgreek{δάμοι}, and the monophthongized dative \textgreek{δάμυ}. Additionally, the data include 4 instances of the genitive singular \textgreek{γᾶς} and 3 of the accusative \textgreek{γᾶν}. Finally, the set includes no \textgreek{η}-type tokens originating from Tanagra. As such, the data suggest city-wide retention of the long \textgreek{ᾱ}. To corroborate this, the author performed Fisher's exact test on the Athenian and Tanagran data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Tanagran data of the 4\super{th} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Tanagran data of the 4\super{th} cent. BCE.

The test resulted in a *p*-value of \< 2.2 x 10^-16^. Therefore, the author rejects the null hypothesis and asserts that the data show a clear relationship between the origin of an inscription and the data type of the tokens therein. Therefore, Tanagra belonged on the Megarian side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 4\super{th} cent. BCE.

```{r 4th_cen_h_tanagra_fisher}
#| include: false

XT.h.4.Tanagra <-
  stats::xtabs( ~ data_type + location, data = Data.h.4.Tanagra)
# Fisher's Exact Test for Tanagra Data
fisher.test(XT.h.4.Tanagra)
```

### 3\super{rd} cent. BCE

```{r Data.h.3}
Data.h.3 <-
  Data.h[Data.h$analysis_cen == 300, ] %>% # . . oi for 4th cen
  droplevels()
```

```{r Count.h.3}
Count.h.3 <- nrow(Data.h.3)
Count.h.3.Athens <- Data.h.3[Data.h.3$location == "Athens",] %>% nrow()
Count.h.3.Oropos <- Data.h.3[Data.h.3$location == "Oropos",] %>% nrow()
Count.h.3.Eleusis <- Data.h.3[Data.h.3$location == "Eleusis",] %>% nrow()
Count.h.3.Megara <- Data.h.3[Data.h.3$location == "Megara",] %>% nrow()
Count.h.3.Aigosthena <- Data.h.3[Data.h.3$location == "Aigosthena",] %>% nrow()
Count.h.3.Pagai <- Data.h.3[Data.h.3$location == "Pagai",] %>% nrow()
Count.h.3.Tanagra <- Data.h.3[Data.h.3$location == "Tanagra",] %>% nrow()
```

The data for the 3\super{rd} cent. BCE shows that the \textgreek{η}/\textgreek{ᾱ} isogloss line remained in the same position as in the 4\super{th} cent. The set contains `r Count.h.3` tokens from across all 7 target cities: Oropos (n=`r Count.h.3.Oropos`), Eleusis (n=`r Count.h.3.Eleusis`), Megara (n=`r Count.h.3.Megara`), Athens (n=`r Count.h.3.Athens`), Tanagra (n=`r Count.h.3.Tanagra`), Aigosthena (n=`r Count.h.3.Aigosthena`), and Pagai (n=`r Count.h.3.Pagai`). 30 \textgreek{ᾱ}-type tokens appear between Oropos and Eleusis, but otherwise the \textgreek{η}- and \textgreek{ᾱ}-type tokens remain neatly on their respective sides of the isogloss line.

```{r Visualize_3rd_cen_h_tokens_by_city, fig.pos='p'}
#| label: fig_3rd_cen_h_tokens
#| fig-cap: "Third Century BCE Token Counts by City"

# Visualize location counts
Data.h.3 %>%
  ggplot(aes(x = forcats::fct_infreq(location), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Location", y = "Token Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r 3rd_cen_h_athens, fig.pos='p'}
#| label: fig_3rd_h_athens
#| fig-cap: "Third Century BCE Athenian Token Counts"

# Visualize tokens in Athens
Data.h.3[Data.h.3$location == "Athens",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

#### Athens

The `r Count.h.3.Athens` Athenian tokens show unanimous adoption of the raised \textgreek{η} instead of the long \textgreek{ᾱ} in both root syllables and \textipa{\=a}-stem singular endings. Without any \textgreek{ᾱ}-type tokens, the data establish Athens as representing the \textgreek{η} side of the \textgreek{η}/\textgreek{ᾱ} isogloss and as the main point of comparison for statistical tests.

First, the data include 21 root-syllable tokens: 19 instances of \textgreek{δῆμος} and 2 of \textgreek{στήλη}. This list includes 7 occurrences of the genitive singular \textgreek{δήμου} and 6 of the dative \textgreek{δήμωι}, along with 1 each of the nominative proper nouns \textgreek{Δημοσθένης} 'Might of the Deme' and \textgreek{Δημοκλῆς} 'Glory of the Deme,' each with the root vowel \textgreek{η}. Furthermore, the 2 instances of \textgreek{στήλη}, 1 genitive singular \textgreek{στήλης} and 1 accusative \textgreek{στήλην}, show the same vocalic quality, indicating that Athenians retained the raised vowel \textgreek{η} in root syllables through the 3\super{rd} cent. BCE.

Next, the Athenian data include 19 \textipa{\=a}-stem tokens, all in forms of \textgreek{βουλή}. The list comprises 9 genitive singulars \textgreek{βουλῆς}, 7 nominatives \textgreek{βουλή}, 2 accusatives \textgreek{βουλήν}, and 1 dative \textgreek{βουλῆι}. As with the root-syllable tokens, each instance of \textgreek{βουλή} shows the continued retention of the raised vowel \textgreek{η} into the 2\super{nd} cent. BCE.

Finally, the Athenian set contains no \textgreek{ᾱ}-type tokens. Therefore, the data categorize Athens as having retained the innovative vowel \textgreek{η} and establish the city as the main point of comparison for later testing.

#### Eleusis

Notwithstanding the city's 3 \textgreek{ᾱ}-type tokens, the `r Count.h.3.Eleusis` Eleusinean tokens show a general retention of the innovative \textgreek{η}. Given that these tokens originate from the same inscription, which otherwise prefers \textgreek{η}-type tokens, they cannot represent a city-wide dialect feature. Besides these, the set shows general retention of the raised \textgreek{η} in root syllables and singular \textipa{\=a}-stem endings.

First, the 3 \textgreek{ᾱ}-type tokens cannot provide sufficient grounds to argue for any significant reversal to or adoption of \textgreek{ᾱ}-type forms within Eleusis. These tokens come exclusively from IG II^2^ 1,299 (PH# 3,513): a long honorific praising 58 individuals, including, crucially, 23 foreigners in the city. The text of line 93 on column 1 explicitly mentions these men as foreigners: \textgreek{καὶ} \textgreek{οἵδε} \textgreek{ξένων} \textgreek{συνανέθηκαν} 'these men from among the foreigners helped in dedicating this.' Each Eleusinean \textgreek{ᾱ}-type token appears in this list, along with an ethnic epithet denoting their home city. With each individual listed as a foreigner, both categorically before their list and specifically with their epithet, the spelling used for their names must represent the pronunciation used in their native dialect.

Furthermore, the inscription contains 14 other serviceable tokens. 1 is the \textgreek{η}-type genitive singular \textgreek{στήλης}, while the other 13 are forms of \textgreek{δῆμος}. Every token collected from this inscription that occurs outside of a proper noun shows the raised vowel \textgreek{η}, following all other Eleusinean inscriptions from this century. As the names contrast with the rest of the inscription in question, the long \textgreek{ᾱ} cannot represent even a feature of the author's pronunciation. Therefore, these three \textgreek{ᾱ}-type tokens cannot provide evidence for a city-wide shift toward \textgreek{ᾱ}-type forms.

```{=tex}
\begin{table}[htbp]
\centering
\caption{Tokens from IG II\textsuperscript{2} 1,299 (PH\# 3,513)}
\label{tab:eleusis_ph3513}
\begin{tabular}{@{}lll@{}}
\toprule
Token                   &      Type     & $n$ \\ \midrule
\textgreek{στηλης}      & \textgreek{η} & 1   \\
\textgreek{δημου}       & \textgreek{η} & 4   \\
\textgreek{δημον}       & \textgreek{η} & 3   \\
\textgreek{δημος}       & \textgreek{η} & 4   \\
\textgreek{δημωι}       & \textgreek{η} & 1   \\
\textgreek{δημοκρατου}  & \textgreek{η} & 1   \\
\textgreek{αριστοδαμος} & \textgreek{ᾱ} & 2   \\
\textgreek{καλλιδαμος}  & \textgreek{ᾱ} & 1   \\ \bottomrule
\end{tabular}
\end{table}
```
Besides the outlying \textgreek{ᾱ}-type tokens, the data contain 48 occurrences of \textgreek{δῆμος} along with 2 of \textgreek{στήλη}, pointing to general retention of the raised vowel \textgreek{η} in the local dialect. These tokens include 10 instances of the genitive singular \textgreek{δήμου} and 9 of the accusative \textgreek{δῆμον}, 2 of the genitive \textgreek{στήλης}, and various secondary forms such as the verb \textgreek{δεδημάρχηκεν} 'has lead the city' and the name \textgreek{Δημοκλῆς} 'Glory of the Deme.' Each instance of \textgreek{δῆμος} and \textgreek{στήλη} contain the raised vowel \textgreek{η}, showing its retention through the 3\super{rd} cent. BCE.

Finally, the data include 14 instances of \textgreek{η} in \textipa{\=a}-stem endings. This includes the 1 accusative singular \textgreek{γῆν} and 13 forms of \textgreek{βουλή}. The list comprises 7 occurrences of the genitive singular \textgreek{βουλῆς}, 4 of the nominative \textgreek{βουλή}, and 2 of the accusative \textgreek{βουλήν}. Each token includes the vowel \textgreek{η} in place of \textgreek{ᾱ}, further indicating its retention through the century. Therefore, the data show that Eleusis generally maintained the raised vowel and belonged on the Athenian side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 3\super{rd} cent. BCE.

```{r 3rd_cen_h_eleusis}
#| label: fig_3rd_h_eleusis
#| fig-cap: "Third Century BCE Eleusinean Token Counts"

# Visualize tokens in Eleusis
Data.h.3[Data.h.3$location == "Eleusis",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

#### Oropos

```{r data_h_3_oropos}
Data.h.3.Oropos <- Data.h.3 %>%
  filter(str_detect(location, "(Athens|Oropos)") == T) %>%
  droplevels()
```

Notwithstanding 28 \textgreek{ᾱ}-type tokens, the `r Count.h.3.Oropos` Oropian tokens show a general retention of the raised \textgreek{η} over the inherited long \textgreek{ᾱ}. The widespread preference for the innovative \textgreek{η} in both root syllables and \textipa{\=a}-stem endings shows that \textgreek{ᾱ}-type inscriptions, though internally consistent, do not represent a city-wide isogloss.

The inscriptional data show a clear preference for the raised vowel \textgreek{η} in the root syllables of \textgreek{δῆμος} and \textgreek{στήλη}. First, all 9 instances of \textgreek{στήλη} contain the vowel \textgreek{η} in place of \textgreek{ᾱ}. This list comprises the 6 occurrences of the dative singular \textgreek{στήληι}, 2 of the accusative \textgreek{στήλην}, and 1 of the nominative \textgreek{στήλη}. Furthermore, of the 163 occurrences of \textgreek{δῆμος}, 147 contain the raised vowel, as seen in the dative singular \textgreek{δήμωι} (n=113), the accusative \textgreek{δῆμον} (n=5), and the nominative proper noun \textgreek{Ἐχέδημος} (n=1). Meanwhile, the low vowel \textgreek{ᾱ} appears in only 16 tokens. The dative singular \textgreek{δάμοι} occurs 4 times, and the monophthongized dative \textgreek{δάμυ} 3 times. However, the remaining 9 tokens are proper nouns, only 2 of which cooccur with other collected \textgreek{ᾱ}-type tokens. These are the nominative \textgreek{Δαμόκριτος} 'Chosen One of the Deme' in the phrase \textgreek{Δαμόκριτος Τιμογένιος Ὀωρώπιος ἔλεξε} 'the Oropian Damokritos Son of Timogenos declared it' from IG VII 393 (PH# 143,870) and the genitive \textgreek{Δαμοφίλω} 'Friend of the Deme' in the formula \textgreek{Δαμοφίλω ἄρχοντος} 'during the Archonship of Damophilo' from IG VII 352 (PH# 143,829).

This general distribution of \textgreek{ᾱ}-type tokens in inscriptions otherwise preferring \textgreek{η}-type forms suggests that the preference for the low vowel was not a general feature of the local dialect. Nevertheless, that these two inscriptions consistently prefer forms with the low vowel \textgreek{ᾱ} and that IG VII 393 (PH# 143,870) mentions explicitly that the motion's mover was an Oropian raises the possibility that a significant portion of the local population did retain the low vowel, warranting further testing.

However, the extant forms of \textgreek{γῆ} and \textgreek{βουλή} also point to a more general preference for the raised vowel \textgreek{η}. Of the 220 instances of \textgreek{γῆ}, 208 show the raised vowel, as in the genitive singular \textgreek{γῆς} (n=107) and the accusative \textgreek{γῆν} (n=101). The list also includes 12 instances of \textgreek{γᾶ}, 6 times each as the genitive singular \textgreek{γᾶς} and the accusative \textgreek{γᾶν}. Though they represent the decided minority of \textgreek{γῆ}-type tokens, their presence also warrants statistical testing.

Furthermore, 31 of the 32 instances of \textgreek{βουλή} show uniform retention of the raised vowel. This includes 27 instances of the dative singular \textgreek{βουλῆι} and 2 of the accusative \textgreek{βουλήν}. Note that this includes 2 aberrant forms captured as \textgreek{βουλη}. The first, from the combined text IG VII 388 + 446 + 494 (PH# 143,865 + 143,923 + 143,971), appears in the phrase \textgreek{τῇ} \textgreek{βουλῇ} \textgreek{καὶ} \textgreek{τῶι} \textgreek{δάμωι}, where the dative singular \textgreek{βουλῆι} has been transcribed with an iota subscript rather than the PHI's typical adscript. Due to the scraper removing all subscripts, the form \textgreek{βουλῇ} has been captured as \textgreek{βουλη}. The second, from IG VII 405 (PH# 143,882), appears in the phrase \textgreek{ἐμ βουλῆ Θυΐο{\lbrk}}. Unfortunately, the inscription is highly fragmentary and lacks any other examples of \textgreek{αι} or \textgreek{η}, leaving it unclear whether the dative represents \textgreek{βουλῆι}, displaying the Attic-Ionic vowel shift with a simple misspelling, or a Boiotian monophthongization of the dative \textgreek{βουλᾶι}.

However, the preceding line contains the genitive article \textgreek{τοῦ} rather than the Boiotian \textgreek{τῶ} seen in the Tanagran inscription DGE 462.A \textgreek{τῶ ἱαρῶ} 'of the temple' and \textgreek{τῶ ἀρχιτέκτονος} 'of the architect,' which appears alongside the monophthongized -\textgreek{η} in \textgreek{τῆ ἀρχῆ} 'for the office' and \textgreek{ἐν ἀνγραφῆ} 'on the register.' Concurrent \textgreek{ου} \> \textgreek{ω} and \textgreek{αι} \> \textgreek{η} shifts also appear in the Oropian inscription DGE 449, from the same century, in the phrase \textgreek{τοῖς λοιποῖς προξένοις κὴ εὐεργέτηις τῶ κοινῶ Βοιωτῶν} 'for the remaining public friends and benefactors of the league of Boiotians.' While the lack of a comparable \textgreek{τοῦ} \> \textgreek{τῶ} shift in IG VII 405 (PH# 143,882) does not necessarily exclude the possibility that \textgreek{βουλῆ} resulted from an older form \textgreek{βουλᾶι}, the absence of this shift in the context of other coeval inscriptions from the region raises the likelihood that it represents a misspelling of \textgreek{βουλῆι}. Therefore, the form has remained in the data as the second instance of \textgreek{βουλη}.

With the 2 aberrant forms accounted for, the instances of \textgreek{βουλή} all point to a general preference for the raised vowel \textgreek{η}. Additionally, with the vast majority of occurrences of \textgreek{γῆ} showing the same preference, the data suggest that the population of Oropos generally accepted the innovative \textgreek{η} over the older long \textgreek{ᾱ}. To corroborate this, the author performed Fisher's exact test on the Athenian and Oropian data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Oropian data of the 3\super{rd} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Oropian data of the 3\super{rd} cent. BCE.

The test resulted in a *p*-value of 0.156: therefore, the author fails to reject the null hypothesis and cannot assert that, per the data, the distribution of \textgreek{ᾱ}-type tokens in Athenian and Oropian inscriptions results from a text's origin in either city. As such, the data indicates a general preference in the city for \textgreek{η}-type forms, demonstrating that the city belonged on the Athenian side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 3\super{rd} cent. BCE.

```{r 3rd_cen_h_oropos}
#| label: fig_3rd_h_oropos
#| fig-cap: "Selected Third Century BCE Oropian Token Counts"

# Visualize tokens in Oropos
Data.h.3.Oropos[Data.h.3.Oropos$location == "Oropos",] %>%
  group_by(token) %>%
  filter(n() > 1) %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(subtitle = "Tokens Occuring More than Once",
       x = "Token",
       y = "Count",
       caption = "Chart displays 405 of 424 Oropian tokens.") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r test_h_3_oropos}
#| include: false

# Contingency Table of Oropian Data
XT.h.3.Oropos <-
  stats::xtabs(~ data_type + location, data = Data.h.3.Oropos)

# Proportion Table of Oropian Data; rows add up to 1
# Prop.h.3.Oropos <- prop.table(XT.h.3.Oropos, 1)

# Fisher's Exact Test for Oropian Data
fisher.test(XT.h.3.Oropos) # . . . . Oropian fisher's exact test
```

#### Megara

```{r data_h_3_megara}
Data.h.3.Megara <- Data.h.3 %>%
  filter(str_detect(location, "(Athens|Megara)") == T) %>%
  droplevels()
```

The `r Count.h.3.Megara` inscriptions from Megara universally show the inheritance of the long \textgreek{ᾱ}, both in root syllables and in \textipa{\=a}-stem endings. First, the word \textgreek{δᾶμος} appears 35 times, primarily in the dative singular \textgreek{δάμωι} (n=16) and the genitive \textgreek{δάμου} (n=9), but also 10 times in derived forms. For example, the proper noun \textgreek{Εὔδαμος} appears in 3 separate inscriptions, once in the nominative and twice in the genitive as \textgreek{Εὐδάμου}, as does the nominative plural \textgreek{δαμιοργοί} 'magistrates' preceding a list of dedicators setting up an offering to Aphrodite. Furthermore, the form \textgreek{στ\'ᾱλᾱ} appears 6 times: 3 times in the accusative \textgreek{στάλαγ}, with the final \textgreek{ν} assimilated to a following velar consonant, twice in the plain accusative \textgreek{στάλαν}, and once in the dative \textgreek{στάλαι}. On each occasion, the root consistently retains the inherited long vowel \textgreek{ᾱ}.

In addition, the word \textgreek{βουλ\'ᾱ} with a singular \textipa{\=a}-stem ending occurs 14 times in the Megarian set. 12 instances are of the dative singular \textgreek{βουλᾶι}, and 2 of the genitive \textgreek{βουλᾶς}. Each contains an \textipa{\=a}-stem ending with the inherited long \textgreek{ᾱ}, showing the same retention of the inherited vowel within the city.

```{r 3rd_cen_h_megara, fig.pos='p'}
#| label: fig_3rd_h_megara
#| fig-cap: "Third Century BCE Megarian Token Counts"

# Visualize tokens in Megara
Data.h.3[Data.h.3$location == "Megara",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{=tex}
\begin{table}[p]
\centering
\caption{Third Century BCE Tokens from Pagai}
\label{tab:data_h_3_pagai}
\begin{tabular}{@{}llllll@{}}
\toprule
Inscription & PH\#     & Location & Analysis Cent. & Token                 & Data Type     \\ \midrule
IG VII 189  & 143,666 & Pagai    & 3             & \textgreek{στάλαν}    & \textgreek{α} \\
IG VII 189  & 143,666 & Pagai    & 3             & \textgreek{Δαμοτίωνα} & \textgreek{α} \\ \bottomrule
\end{tabular}
\end{table}
```

This local preference for \textgreek{ᾱ} over \textgreek{η} in both roots and \textipa{\=a}-stem endings suggests that pronunciations with a long \textgreek{ᾱ} were a general feature of the local dialect. To corroborate this, the author performed Fisher's exact test on the Athenian and Megarian data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Megarian data of the 3\super{rd} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Megarian data of the 3\super{rd} cent. BCE.

The test returned a *p*-value of 2.2 x 10^-16^. Therefore, the author rejects the null hypothesis and asserts that the distribution of data types between inscriptions from Megara and Athens is related to the city of origin for such inscriptions. As such, the data indicate that Megara belonged on the \textgreek{ᾱ} side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 3\super{rd} cent. BCE.

```{r test_h_3_megara}
#| include: false

# Contingency Table of Megarian Data
XT.h.3.Megara <-
  stats::xtabs(~ data_type + location, data = Data.h.3.Megara)

# Proportion Table of Megarian Data; rows add up to 1
# Prop.h.3.Megara <- prop.table(XT.h.3.Megara, 1)

# Fisher's Exact Test for Megarian Data
fisher.test(XT.h.3.Megara) # . . . . Megarian fisher's exact test
```

#### Pagai

```{r data_h_3_pagai}
Data.h.3.Pagai <- Data.h.3 %>%
  filter(str_detect(location, "(Athens|Pagai)") == T) %>%
  droplevels()
```

The data from Pagai include `r Count.h.3.Pagai` tokens. These are the accusative singular \textgreek{στάλαν} and the nominative \textgreek{Δαμοτίων} 'Of the Deme.' In each instance, the root maintains the inherited long vowel \textgreek{ᾱ}, suggesting a general preference for the long \textgreek{ᾱ} over \textgreek{η} in the city. To corroborate this, the author performed Fisher's exact test on the Athenian and Pagaian data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Pagaian data of the 3\super{rd} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Pagaian data of the 3\super{rd} cent. BCE.

The test resulted in a *p*-value of 0.001161. Therefore, the author rejects the null hypothesis and asserts that the distribution of data types among inscriptions between Athens and Pagai relates to an inscription's city of origin. For this reason, and given the lack of any \textgreek{η}-type evidence from the city, the data indicate that Pagai belonged on the Megarian side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 3\super{rd} cent. BCE

```{r test_h_3_pagai}
#| include: false

# Contingency Table of Pagaian Data
XT.h.3.Pagai <-
  stats::xtabs(~ data_type + location, data = Data.h.3.Pagai)

# Proportion Table of Pagaian Data; rows add up to 1
# Prop.h.3.Pagai <- prop.table(XT.h.3.Pagai, 1)

# Fisher's Exact Test for Pagaian Data
fisher.test(XT.h.3.Pagai) # . . . . Pagaian fisher's exact test
```

#### Aigosthena

```{r data_h_3_aigosthena}
Data.h.3.Aigosthena <- Data.h.3 %>%
  filter(str_detect(location, "(Athens|Aigosthena)") == T) %>%
  droplevels()
```

Similarly, the `r Count.h.3.Aigosthena` tokens from Aigosthena suggest a general retention of the low vowel \textgreek{ᾱ} in root syllables and \textipa{\=a}-stem endings. The set includes 6 instances of \textgreek{δᾶμος}: 3 times in the dative singular \textgreek{δάμωι} and once in the shortened dative \textgreek{δάμοι}, as well as once each in the nominative \textgreek{Χαρίδαμος} 'Grace of the Deme' and the genitive \textgreek{Δαμοκλέος} 'Glory of the Deme.' Additionally, the set includes 2 occurrences of the accusative singular \textgreek{στάλαν} and 1 of the accusative \textgreek{γᾶν}. Overall, the Aigosthenian data exclusively contain the long vowel \textgreek{ᾱ} in the targeted root and ending segments, suggesting general retention in the city. To corroborate this, the author performed Fisher's exact test on the Athenian and Aigosthenain data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Aigosthenian data of the 3\super{rd} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Aigosthenian data of the 3\super{rd} cent. BCE.

The test resulted in a *p*-value of 4.867 x 10^-10^ Therefore, the author rejects the null hypothesis and asserts that the distribution of data types between tokens from Athens and Aigosthena relates to the city of origin for an inscription. As such, per the collected data, Aigosthena belonged to the Megarian side of the \textgreek{η}/\textgreek{ᾱ} isogloss during the 3\super{rd} cent. BCE.

```{r 3rd_cen_h_aigosthena}
#| label: fig_3rd_h_aigosthena
#| fig-cap: "Third Century BCE Aigosthenian Token Counts"

Data.h.3[Data.h.3$location == "Aigosthena",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r test_h_3_aigosthena}
#| include: false

# Contingency Table of Aigosthenian Data
XT.h.3.Aigosthena <-
  stats::xtabs(~ data_type + location, data = Data.h.3.Aigosthena)

# Proportion Table of Aigosthenian Data; rows add up to 1
# Prop.h.3.Aigosthena <- prop.table(XT.h.3.Aigosthena, 1)

# Fisher's Exact Test for Aigosthenian Data
fisher.test(XT.h.3.Aigosthena) # . . . . Aigosthenian fisher's exact test
```

#### Tanagra

```{r data_h_3_tanagra}
Data.h.3.Tanagra <- Data.h.3 %>%
  filter(str_detect(location, "(Athens|Tanagra)") == T) %>%
  droplevels()
```

The `r Count.h.3.Tanagra` tokens from Tanagra indicate that the city's population retained the long vowel \textgreek{ᾱ} in root syllables and in \textipa{\=a}-stem endings. The set includes 7 occurrences of  \textgreek{δᾶμος} and 11 of \textgreek{γᾶ}, demonstrating the retention in both contexts.

First, \textgreek{δᾶμος} appears in the dative singular \textgreek{δάμοι} (n=4), the dative monophthongized \textgreek{δάμυ} (n=2), and the genitive proper noun \textgreek{Δαμοφίλω} 'Friend of the Deme' (n=1). In each instance, the root contains the inherited long vowel \textgreek{ᾱ}, suggesting general conservation in root syllables. Furthermore, the list contains 6 instances of the genitive singular \textgreek{γᾶς} and 5 of the accusative \textgreek{γᾶν}, suggesting the same conservation in \textipa{\=a}-stem endings. To corroborate this, the author performed Fisher's exact test on the Athenian and Tanagran data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Tanagran data of the 3\super{rd} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Tanagran data of the 3\super{rd} cent. BCE.

The test returned a *p*-value of 2.222 x 10^-15^. Therefore, the author rejects the null hypothesis and asserts that the distribution of \textgreek{η}-type and \textgreek{ᾱ}-type tokens between Athens and Tanagra relates to the origins of the inscriptions collected. Following this, the data indicate that Tanagra belonged on the Megarian side of the \textgreek{η}/\textgreek{ᾱ} isogloss during the 3\super{rd} cent. BCE.

```{r 3rd_cen_h_tanagra}
#| label: fig_3rd_h_tanagra
#| fig-cap: "Third Century BCE Tanagran Token Counts"

Data.h.3[Data.h.3$location == "Tanagra",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r test_h_3_tanagra}
#| include: false

# Contingency Table of Tanagran Data
XT.h.3.Tanagra <-
  stats::xtabs(~ data_type + location, data = Data.h.3.Tanagra)

# Proportion Table of Tanagran Data; rows add up to 1
# Prop.h.3.Tanagra <- prop.table(XT.h.3.Tanagra, 1)

# Fisher's Exact Test for Tanagran Data
fisher.test(XT.h.3.Tanagra) # . . . . Tanagran fisher's exact test
```

### 2\super{nd} cent. BCE

```{r data_h_2}
Data.h.2 <-
  Data.h[Data.h$analysis_cen == 200, ] %>% # . . oi for 2nd cen
  droplevels()
```

```{r Count_h_2}
Count.h.2 <- nrow(Data.h.2)
Count.h.2.Athens <- Data.h.2[Data.h.2$location == "Athens",] %>% nrow()
Count.h.2.Oropos <- Data.h.2[Data.h.2$location == "Oropos",] %>% nrow()
Count.h.2.Eleusis <- Data.h.2[Data.h.2$location == "Eleusis",] %>% nrow()
Count.h.2.Megara <- Data.h.2[Data.h.2$location == "Megara",] %>% nrow()
Count.h.2.Aigosthena <- Data.h.2[Data.h.2$location == "Aigosthena",] %>% nrow()
Count.h.2.Pagai <- Data.h.2[Data.h.2$location == "Pagai",] %>% nrow()
Count.h.2.Tanagra <- Data.h.2[Data.h.2$location == "Tanagra",] %>% nrow()
```

The data for the 2\super{nd} cent. BCE shows that the isogloss line remained stable in dividing Attica and Oropia from Megaris and the rest of Boiotia. The set contains 127 tokens from 5 cities: Athens (n=`r Count.h.2.Athens`), Oropos (n=`r Count.h.2.Oropos`), Eleusis (n=`r Count.h.2.Eleusis`), Megara (n=`r Count.h.2.Megara`), and Aigosthena (n=`r Count.h.2.Aigosthena`). The \textgreek{η}- and \textgreek{ᾱ}-type tokens divide with near exclusivity across the isogloss line, with only 3 \textgreek{ᾱ}-type tokens appearing on the \textgreek{η} side.

```{r vis_data_h_2}
#| fig-cap: "Second Century BCE Token Counts by City"
Data.h.2 %>%
  ggplot(aes(x = forcats::fct_infreq(location), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Location", y = "Token Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

#### Athens

First, the `r Count.h.2.Athens` Athenian tokens indicate a general preference for the raised vowel \textgreek{η} in both root syllables, as seen in forms of \textgreek{δῆμος} and \textgreek{στήλη}, as well as \textipa{\=a}-stem endings, evidenced by forms of \textgreek{βουλή} and \textgreek{γῆ}.

The set from Athens contains 36 \textgreek{η}-type instances of \textgreek{δῆμος} and \textgreek{στήλη}, which together show near exclusive use of the raised vowel. These include 13 instances of the genitive singular \textgreek{δήμου}, 6 of the dative \textgreek{δήμωι} and 4 of the accusative \textgreek{δῆμον} alongside 5 of the genitive \textgreek{στήλης} and 2 of the accusative \textgreek{στήλην}. Although 1 \textgreek{ᾱ}-type token appears in the set, it does not constitute sufficient evidence for a broader Athenian dialectal feature.

The proper noun \textgreek{Δαμόκριτος} 'Chosen One of the Deme' occurs in IG II^2^ 913 (PH# 3,133) and contains the low root vowel \textgreek{ᾱ}. The name appears without any following patronym or epithet, denying us the ability to definitively classify the individual as a visitor, a resident expatriate, or even a native Athenian born to immigrant parents. Nevertheless, the form \textgreek{Δαμ}- directly contradicts the remaining text, which elsewhere shows \textgreek{δῆμος} (n=3) with the raised vowel. The genitive singular \textgreek{δήμου} occurs twice, along with the name \textgreek{Δημοφάνης}. While \textgreek{Δημοφάνης} 'Image of the Deme' faces the same evidentiary issues as \textgreek{Δαμόκριτος}, that the majority of \textgreek{δημ}- tokens prefer the raised vowel over the low vowel suggests the author generally accepted the Attic-Ionic vowel shift. As such, the likelihood remains low that the low vowel in \textgreek{Δαμόκριτος} represents a feature of the author's dialect.

Additionally, the data indicate the continued preference for the raised vowel in \textipa{\=a}-stem endings. The set includes 20 instances of \textgreek{βουλή} in 3 cases: 11 of the nominative \textgreek{βουλή}, 7 of the genitive \textgreek{βουλῆς}, and 2 of the dative \textgreek{βουλῆι}. The set contains no \textgreek{ᾱ}-type \textipa{\=a}-stem endings, showing a general retention of the raised vowel \textgreek{η}.

Overall, the data contains only 1 \textgreek{ᾱ}-type token against 56 \textgreek{η}-types. The \textgreek{ᾱ}-type token contradicts the general preference for \textgreek{η}-types in the same inscription, suggesting a foreign affectation. With no inscriptions containing exclusively \textgreek{ᾱ}-type data, the set indicates the continued preference for the raised vowel \textgreek{η} over the low vowel \textgreek{ᾱ} in Athens during the 2\super{nd} cent. BCE. As such, Athens remains the primary city of comparison for statistical tests.

```{r vis_data_h_2_athens, fig.pos='p'}
#| fig-cap: "Second Century BCE Athenian Token Counts"

Data.h.2[Data.h.2$location == "Athens",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r vis_data_h_2_eleusis, fig.pos='p'}
#| fig-cap: "Second Century BCE Eleusinean Token Counts"

Data.h.2[Data.h.2$location == "Eleusis",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

#### Eleusis

The 22 Eleusinean tokens show the same preference for the raised vowel \textgreek{η} over the low vowel \textgreek{ᾱ}, both in the root syllables of \textgreek{δῆμος} and \textgreek{στήλη}, as well as the \textipa{\=a}-stem endings of \textgreek{βουλή}.

The data contain 17 occurrences \textgreek{δῆμος} (n=14) and \textgreek{στήλη} (n=3). Among these are 2 instances each of the genitive singular \textgreek{δήμου}, the nominative \textgreek{δῆμος}, and the proper noun \textgreek{Δήμαρχος} 'Leader of the Deme.' In addition, the set includes 2 instances of the genitive singular \textgreek{στήλης} and 1 of the accusative plural \textgreek{στήλας}. These stems exclusively contain the raised vowel, indicating its retention in root syllables.

Furthermore, forms of \textgreek{βουλή} appear 5 times in the set. This includes 2 instances of the nominative singular \textgreek{βουλή}, 2 of the genitive \textgreek{βουλής}, and 1 of the accusative \textgreek{βουλήν}. These forms contain only the raised vowel, indicating its retention in \textipa{\=a}-stem endings.

Finally, the data contain no Eleusinean \textgreek{ᾱ}-type tokens. Therefore, they indicate a continued acceptance of the raised vowel \textgreek{η} in the 2\super{nd} cent. BCE.

#### Oropos

```{r data_h_2_oropos}
Data.h.2.Oropos <- Data.h.2 %>%
  filter(str_detect(location, "(Athens|Oropos)") == T) %>%
  droplevels()
```

The 31 Oropian tokens show a majority acceptance of the raised vowel \textgreek{η} in both root syllables and \textipa{\=a}-stem endings. Besides the majority of \textgreek{η}-type occurrences of \textgreek{δῆμος} and \textgreek{γῆ}, the set contains 1 \textgreek{ᾱ}-type token. However, this instance does not sufficiently evidence a general dialectal feature of the city.

The \textgreek{η}-type tokens comprise 1 instance of the genitive singular \textgreek{γῆς} along with 29 \textgreek{η}-type occurrences of \textgreek{δῆμος}. For example, the accusative singular \textgreek{δῆμον} appears 7 times and the dative \textgreek{δήμωι} 5 times, along with 3 instances of the genitive \textgreek{Ἀλεξιδήμου} 'Defender of the Deme,' 2 of \textgreek{Εὐθυδήμου} 'Whose Deme is Honest,' and another 2 of \textgreek{Δημογένου} 'Kin of the Deme.' However, \textgreek{δᾶμος} occurs once in the set in the proper noun \textgreek{Δαμάρχου} from IG VII 3,498 (PH# 147,019), raising the possibility that it reflects a local dialectal variant. The text contains an inventory of silver dedications given to the sanctuary at Amphiaraos, such as ln. 14 \textgreek{σκάφιον Δημάρχου \d{ϙ} ζ}ʹ '96 bowl(s) from Demarchos' and ln. 24 \textgreek{ἄλλος Δημαινέτου ρ}ʹ '90 other (cups) from Demainetos.' As part of the formulaic opening, \textgreek{Δαμάρχου} appears as a patronym in the phrase \textgreek{ἱεραρχούντων Ἱεροκλέους τοῦ Δαμάρχου ...} 'during the high-priesthood of Hierokles son of Damarchos ...' This does not necessarily require that \textgreek{δᾶμος} appeared in a local dialect; however, it places the \textgreek{ᾱ}-type token within 1 generation of a standing Oropian religious official. As such, its appearance warrants further testing. In order to determine to what degree this form suggests local dialectal variation, the author performed Fisher's exact test on the Athenian and Oropian data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Oropian data of the 2\super{nd} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Oropian data of the 2\super{nd} cent. BCE.

The test resulted in a *p*-value of 1, indicating that the distribution of collected \textgreek{ᾱ}-type and \textgreek{η}-type tokens between Athens and Oropos does not relate to an inscription's city of origin. As such, the author fails to reject the null hypothesis and asserts that form \textgreek{Δαμάρχου} does not sufficiently evidence a local dialect by its mere presence among the Oropian data. Therefore, with no other \textgreek{ᾱ}-type tokens collected, the data indicates that Oropos belonged on the Athenian side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 2\super{nd} cent. BCE.

```{r vis_data_h_2_oropos}
#| fig-cap: "Second Century BCE Oropian Token Counts"

Data.h.2[Data.h.2$location == "Oropos",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r test_data_h_2_oropos}
#| include: false

# Contingency Table of Oropian Data
XT.h.2.Oropos <-
  stats::xtabs(~ data_type + location, data = Data.h.2.Oropos)

# Proportion Table of Oropian Data; rows add up to 1
# Prop.h.2.Oropos <- prop.table(XT.h.2.Oropos, 1)

# Fisher's Exact Test for Oropian Data
fisher.test(XT.h.2.Oropos) # . . . . Oropian fisher's exact test
```

#### Megara

```{r data_h_2_megara}
Data.h.2.Megara <- Data.h.2 %>%
  filter(str_detect(location, "(Athens|Megara)") == T) %>%
  droplevels()
```

The `r Count.h.2.Megara` Megarian tokens point to continued retention of the low vowel \textgreek{ᾱ} in roots, via forms of \textgreek{δᾶμος} (n=9) and \textgreek{στ\'ᾱλᾱ} (n=2), as well as in \textipa{\=a}-stem endings, as seen in declensions of \textgreek{γᾶ} (n=1) and \textgreek{βουλ\'ᾱ} (n=1).

The set comprises 11 \textgreek{ᾱ}-type occurrences of \textgreek{δᾶμος} and \textgreek{στ\'ᾱλᾱ}. These include 3 instances of the accusative singular \textgreek{δᾶμαν}, 2 of the nominative \textgreek{δᾶμος}, and 2 instances of the accusative singular \textgreek{στάλαν}. Furthermore, the scraper did not collect any \textgreek{η}-type occurrences of \textgreek{δῆμος} or \textgreek{στήλη}, indicating total retention of the low vowel \textgreek{ᾱ} in roots. Additionally, the set contains 1 instance of each of the accusative singulars \textgreek{γᾶν} and \textgreek{βουλάν}, suggesting the same retention in \textipa{\=a}-stem endings. The unanimity of \textgreek{ᾱ}-type data in Megara compared to the vast majority of \textgreek{η}-type data in Athens suggests a relationship between an inscription's city of origin and the type of tokens found therein. To test for a relationship between a token's origin and its data type, the author performed Fisher's exact test on the Athenian and Oropian data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Megarian data of the 2\super{nd} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Megarian data of the 2\super{nd} cent. BCE.

The test resulted in a *p*-value of 2.949 x 10^-13^. Therefore, the author rejects the null hypothesis and asserts that the distribution of \textgreek{η}-type and \textgreek{ᾱ}-type tokens between Athens and Megara relates to the origins of the inscriptions collected. Accordingly, the data indicate that Megara belonged on the \textgreek{ᾱ} side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 2\super{nd} cent. BCE.

```{r vis_data_h_2_megara}
#| fig-cap: "Second Century BCE Megarian Token Counts"

Data.h.2[Data.h.2$location == "Megara",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r test_data_h_2_megara}
#| include: false

# Contingency Table of Megarian Data
XT.h.2.Megara <-
  stats::xtabs(~ data_type + location, data = Data.h.2.Megara)

# Proportion Table of Megarian Data; rows add up to 1
# Prop.h.2.Megara <- prop.table(XT.h.2.Megara, 1)

# Fisher's Exact Test for Megarian Data
fisher.test(XT.h.2.Megara) # . . . . Megarian fisher's exact test
```

#### Aigosthena

```{r data_h_2_aigosthena}
Data.h.2.Aigosthena <- Data.h.2 %>%
  filter(str_detect(location, "(Athens|Aigosthena)") == T) %>%
  droplevels()
```

Finally, the 4 tokens from Aigosthena strongly suggest that the city retained the low vowel \textgreek{ᾱ} in all positions. This set comprises 2 instances of the dative singular \textgreek{δάμωι}, 1 of the accusative \textgreek{δᾶμον}, and 1 of the genitive singular \textgreek{γᾶς}. Furthermore, the lack of any \textgreek{η}-type tokens, when compared to the majority of \textgreek{η}-type tokens in Athens, suggests that an inscription's city of origin relates to the data type found therein.

To test for this relationship, the author performed Fisher's exact test on the Athenian and Oropian data with the following hypotheses:

H0

:   There is no relationship between the variables `location` and `data_type` in the combined Athenian and Aigosthenian data of the 2\super{nd} cent. BCE.

HA

:   There is a relationship between the variables `location` and `data_type` in the combined Athenian and Aigosthenian data of the 2\super{nd} cent. BCE.

The test returned a *p*-value of 9.581 x 10^-06^. Therefore, the author rejects the null hypothesis and asserts that the distribution of \textgreek{η}-type and \textgreek{ᾱ}-type tokens between Athens and Aigosthena relates to the origins of the inscriptions collected. Following this, the data collected indicate that Aigosthena belonged on the Megarian side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 2\super{nd} cent. BCE.

```{r vis_data_h_2_aigosthena}
#| fig-cap: "Second Century BCE Aigosthenian Token Counts"

Data.h.2[Data.h.2$location == "Aigosthena",] %>%
  ggplot(aes(x = forcats::fct_infreq(token), fill = data_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count))) +
  labs(x = "Token", y = "Count") +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values=iso_colors)
```

```{r test_data_h_2_aigosthena}
#| include: false

# Contingency Table of Aigosthenian Data
XT.h.2.Aigosthena <-
  stats::xtabs(~ data_type + location, data = Data.h.2.Aigosthena)

# Proportion Table of Aigosthenian Data; rows add up to 1
# Prop.h.2.Aigosthena <- prop.table(XT.h.2.Aigosthena, 1)

# Fisher's Exact Test for Aigosthenian Data
fisher.test(XT.h.2.Aigosthena) # . . . . Aigosthenian fisher's exact test
```

### Interpretations

Overall, the data present a stable isogloss line between the 4^th^ and 2^nd^ cents.\ BCE. However, evidentiary issues make it challenging to place Pagai, Aigosthena, and Tanagra with any exactness.

The 3^rd^ cent. somewhat alleviates this problem for Pagai and Aigosthena. This century's data set contains observations from all 7 target cities, thereby distinguishing Eleusis, Athens, and Oropos from the rest of central Greece. Given the similarities to the \textgreek{οι}/\textgreek{υ} isogloss -- where the 3^rd^ cent. also included evidence from every city, while the 4^th^ lacked data from Aigosthena and Pagai, and where the Pagaian and Aigosthenian data corresponded with the Megarian data as available -- the author has tentatively placed the two cities on the \textgreek{ᾱ} side of the \textgreek{η}/\textgreek{ᾱ} isogloss line during the 4^th^ and 2^nd^ cents.\ BCE, with the corresponding segments drawn with a dashed line.

However, Tanagra in the 2^nd^ cent. BCE presents further difficulties. In addition to the lack of any inscriptional evidence, this city is the only Boiotian settlement in this study, impeding any efforts to hypothesize on the local dialect's features based on contemporary data. For comparison, Pagai and Aigosthena were historically under the control of Megara, and so the persistence of Megarian Greek features in their local dialects (e.g., nominative articles in \textgreek{τοί} without monophthongization of \textgreek{οι} to \textgreek{υ}, retention of the long low vowel \textgreek{ᾱ}) indicates a shared Doric vernacular. As a result, the Megarian data can provide some insight into the dialect of these cities when their inscriptions do not appear in the data set. However, though Oropos was within the same geographic region, and though it was separated from Athens by Mt. Parnes to the south, the collected data show a majority Attic-Ionic dialect spoken throughout the city between the 4\super{th} and 2\super{nd} cents.\ BCE. Given that no other city's population appeared to speak a Boiotian variety, the 2\super{nd} cent.\ data does not enable any cross-city comparison with Tanagra. Therefore, the placement of this city during the 2^nd^ cent. BCE relies entirely on data from the 3\super{rd} and 4\super{th} cents., which suggest local retention of the long low vowel \textgreek{ᾱ}. The line has been drawn accordingly, with the hypothetical boundaries marked off with a dashed line.

```{r map_4th_cen_h, fig.width=7, fig.height=4, fig.align='center'}
#| fig-cap: "Fourth Century BCE Isogloss Line"

# Layer contour map
contour(elev,
        frame.plot = F,
        col = terrain.colors(n=10),
        xlab = "Latitude",
        ylab = "Longitude")
# Plot cities
points(
  x = cities$x,
  y = cities$y,
  col = "black",
  pch = 19,
  xlim = c(23, 24.5),
  ylim = c(37.5, 38.5)
)

# Define x-points for iso
h_4th_x <- c(
  ((cities[6,1] + cities[7,1])/2)+.05, # Tanagra v Oropos
  (cities[6,1] + cities[2,1])/2, # Tanagra v Eleusis
  (cities[5,1] + cities[2,1])/2, # Aigosthena v Eleusis
  (cities[4,1] + cities[2,1])/2, # Pagai v Eleusis
  (cities[3,1] + cities[2,1])/2, # Megara v Eleusis
  ((cities[3,1] + cities[1,1])/2)-.05  # Megara v Athens
)

# Define y-points for iso
h_4th_y <- c(
  ((cities[6,2] + cities[7,2])/2)+.05, # Tanagra v Oropos
  (cities[6,2] + cities[2,2])/2, # Tanagra v Eleusis
  (cities[5,2] + cities[1,2])/2, # Aigosthena v Eleusis
  (cities[4,2] + cities[1,2])/2, # Pagai v Eleusis
  (cities[3,2] + cities[1,2])/2, # Megara v Eleusis
  ((cities[3,2] + cities[1,2])/2)-.05  # Megara v Athens
)

# Draw solid line (Tanagra)
lines(x = h_4th_x[1:2], y = h_4th_y[1:2], col = "blue", lwd = 2)

# Draw solid line (Megara)
lines(x = h_4th_x[4:6], y = h_4th_y[4:6], col = "blue", lwd = 2)

# Draw dashed line
lines(x = h_4th_x, y = h_4th_y, col = "blue", lwd = 2, lty = 3)

# Label cities
# Moved to end so that iso lines wont overlap text
text(x = cities[-c(2,5,6),]$x,
     y = cities[-c(2,5,6),]$y,
     labels = cities[-c(2,5,6),]$names,
     col = "black",
     pos = 1)
text(x = cities[-c(1,2,3,4,7),]$x,
     y = cities[-c(1,2,3,4,7),]$y,
     labels = cities[-c(1,2,3,4,7),]$names,
     col = "black",
     pos = 3
)
text(x = cities[-c(1,3,4,5,6,7),]$x+.1,
     y = cities[-c(1,3,4,5,6,7),]$y-.01,
     labels = cities[-c(1,3,5,4,6,7),]$names,
     col = "black",
     pos = 3
)
```

```{r map_3rd_cen_h, fig.width=7, fig.height=4, fig.align='center'}
#| fig-cap: "Third Century BCE Isogloss Line"

# Layer contour map
contour(elev,
        frame.plot = F,
        col = terrain.colors(n=10),
        xlab = "Latitude",
        ylab = "Longitude")
# Plot cities
points(
  x = cities$x,
  y = cities$y,
  col = "black",
  pch = 19,
  xlim = c(23, 24.5),
  ylim = c(37.5, 38.5)
)

# Define x-points for iso
h_3rd_x <- c(
  ((cities[6,1] + cities[7,1])/2)+.05, # Tanagra v Oropos
  (cities[6,1] + cities[2,1])/2, # Tanagra v Eleusis
  (cities[5,1] + cities[2,1])/2, # Aigosthena v Eleusis
  (cities[4,1] + cities[2,1])/2, # Pagai v Eleusis
  (cities[3,1] + cities[2,1])/2, # Megara v Eleusis
  ((cities[3,1] + cities[1,1])/2)-.05  # Megara v Athens
)

# Define y-points for iso
h_3rd_y <- c(
  ((cities[6,2] + cities[7,2])/2)+.05, # Tanagra v Oropos
  (cities[6,2] + cities[2,2])/2, # Tanagra v Eleusis
  (cities[5,2] + cities[1,2])/2, # Aigosthena v Eleusis
  (cities[4,2] + cities[1,2])/2, # Pagai v Eleusis
  (cities[3,2] + cities[1,2])/2, # Megara v Eleusis
  ((cities[3,2] + cities[1,2])/2)-.05  # Megara v Athens
)

# Draw iso line
lines(x = h_3rd_x, y = h_3rd_y, col = "blue", lwd = 2)

# Label cities
# Moved to end so that iso lines wont overlap text
text(x = cities[-c(2,5,6),]$x,
     y = cities[-c(2,5,6),]$y,
     labels = cities[-c(2,5,6),]$names,
     col = "black",
     pos = 1)
text(x = cities[-c(1,2,3,4,7),]$x,
     y = cities[-c(1,2,3,4,7),]$y,
     labels = cities[-c(1,2,3,4,7),]$names,
     col = "black",
     pos = 3
)
text(x = cities[-c(1,3,4,5,6,7),]$x+.1,
     y = cities[-c(1,3,4,5,6,7),]$y-.01,
     labels = cities[-c(1,3,5,4,6,7),]$names,
     col = "black",
     pos = 3
)
```

```{r map_2nd_cen_h, fig.width=7, fig.height=4, fig.align='center'}
#| fig-cap: "Second Century BCE Isogloss Line"

# Layer contour map
contour(elev,
        frame.plot = F,
        col = terrain.colors(n=10),
        xlab = "Latitude",
        ylab = "Longitude")
# Plot cities
points(
  x = cities$x,
  y = cities$y,
  col = "black",
  pch = 19,
  xlim = c(23, 24.5),
  ylim = c(37.5, 38.5)
)

# Define x-points for iso
h_2nd_x <- c(
  ((cities[6,1] + cities[7,1])/2)+.05, # Tanagra v Oropos
  (cities[6,1] + cities[2,1])/2, # Tanagra v Eleusis
  (cities[5,1] + cities[2,1])/2, # Aigosthena v Eleusis
  (cities[4,1] + cities[2,1])/2, # Pagai v Eleusis
  (cities[3,1] + cities[2,1])/2, # Megara v Eleusis
  ((cities[3,1] + cities[1,1])/2)-.05  # Megara v Athens
)

# Define y-points for iso
h_2nd_y <- c(
  ((cities[6,2] + cities[7,2])/2)+.05, # Tanagra v Oropos
  (cities[6,2] + cities[2,2])/2, # Tanagra v Eleusis
  (cities[5,2] + cities[1,2])/2, # Aigosthena v Eleusis
  (cities[4,2] + cities[1,2])/2, # Pagai v Eleusis
  (cities[3,2] + cities[1,2])/2, # Megara v Eleusis
  ((cities[3,2] + cities[1,2])/2)-.05  # Megara v Athens
)

# Draw iso line

# Draw solid line (Megara)
lines(x = h_4th_x[3:6], y = h_4th_y[3:6], col = "blue", lwd = 2)

# Draw dashed line
lines(x = h_4th_x, y = h_4th_y, col = "blue", lwd = 2, lty = 3)

# Label cities
# Moved to end so that iso lines wont overlap text
text(x = cities[-c(2,5,6),]$x,
     y = cities[-c(2,5,6),]$y,
     labels = cities[-c(2,5,6),]$names,
     col = "black",
     pos = 1)
text(x = cities[-c(1,2,3,4,7),]$x,
     y = cities[-c(1,2,3,4,7),]$y,
     labels = cities[-c(1,2,3,4,7),]$names,
     col = "black",
     pos = 3
)
text(x = cities[-c(1,3,4,5,6,7),]$x+.1,
     y = cities[-c(1,3,4,5,6,7),]$y-.01,
     labels = cities[-c(1,3,5,4,6,7),]$names,
     col = "black",
     pos = 3
)
```
