\documentclass[./dissertation.tex]{subfiles}
\begin{document}

  \chapter*{Acknowledgments}

  %This will add the acknowledgements back to the toc
  \addcontentsline{toc}{chapter}{Acknowledgments}

I am indebted to Dr.\ Klein, from whom I have learned everything I know of historical linguistics, and Dr.\ Pickett, from whom I have learned everything I know of epigraphy.
\vspace*{1\baselineskip}

{\noindent}I am eternally indebted to Dr.\ Platter, by whose efforts my grasp of the Ancient Greek language has strengthened manyfold. Were it not for his patience and determination, surely I would have succumbed to the \textit{avoirdupois} of my studies.
\vspace*{1\baselineskip}

{\noindent}I am grateful to my family -- parents, siblings, and myriad cousins all -- who never lost faith in me even as my dreams seemed to slip away.
\vspace*{1\baselineskip}

{\noindent}I am grateful to the Kennedys for housing and supporting me through the pandemic, for treating me like family, and for showing me that seeking out an answer can be more valuable than having one.
\vspace*{1\baselineskip}

{\noindent}I am grateful to Jo and Liz, both for their bottomless confidence and their gentle reminders that whoever stands in the path shows you the path.
\vspace*{1\baselineskip}

{\noindent}I am thankful for Rachel, who has shown me nothing but friendship and grace for all these years, and willingly -- even enthusiastically -- endures my endless questions on every topic in marine biology. Few come to mind who would travel so far just for a visit, and I cherish every mile.
\vspace*{1\baselineskip}

{\noindent}I am boundlessly thankful for Brienne, who supported me through every trial. She fortified me when I thought of quitting before attaining my Bachelor's degree, encouraged me with every application, and stood by my side through every struggle in pursuit of my Master's. She supported me through three jobs and four-hour commutes; through outbreaks and furlough; through unemployment, then night shifts, then opening shifts; through business travel, then the choice to give it up in the pursuit of my most quixotic ambitions. The change was terrifying, but what can occur without change?

She is my courage and my temperance, helping me not to shrink with anxiety or act in haste. She fills my every day with happiness and tranquility, and is my life's truest partner.
\vspace*{1\baselineskip}

{\noindent}Most of all, I am grateful to Cher, my curly-haired lap dog, for keeping my feet warm at night and ensuring I wake up on time in the morning. Don't worry: you'll get that FedEx guy someday.

\end{document}
